---
title: "step4_create.70-fold.Rmd"
author: "Adam Festa"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

The purpose of this script is to generate FDR adjusted p-values using the 70 training groups with SNPs & Transcripts

## Load Data 

* Load the starting dataset genereated in step 2 (ordering data)

```{r load.dat}
library(caret);library(parallel)
load("~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/ordered.data.RData")
names(ordered.data)
```

## Create train/test folds

* Create the 70 train/test groups by first setting the seed and then using caret to create 10 replicates of 7-fold cross validation

* Family names of the training and test groups are placed in amatrix

```{r train.test}
set.seed(24826)
all.folds <- createMultiFolds(y=ordered.data$fam.phenos$Volume,times = 10,k = 7)
# Each Rep corresponds to a single 7-fold CV (Rep 1: folds 1-7 rep1; Rep2: folds 1-7 rep2)

train.index.matrix <- do.call(rbind,all.folds)
test.index.matrix <- do.call(rbind,lapply(1:70,function(x) {
  c(1:56)[-c(train.index.matrix[x,])]
}))

train.fam.matrix <- do.call(rbind,lapply(1:70,function(x) {
  as.character(ordered.data$fam.phenos$fam_id)[c(train.index.matrix[x,])]
}))
test.fam.matrix <- do.call(rbind,lapply(1:70,function(x) {
  as.character(ordered.data$fam.phenos$fam_id)[-c(train.index.matrix[x,])]
}))
```


## Conduct anova

* First conduct anova on transcripts for each of the 70 groups.  The biological replicate counts are first log2-transformed and scaled across all counts

```{r anova.txpts}
log.bio.counts <- scale(log2(ordered.data$bio.counts + 1),center = F)

pt <- proc.time()
txpt.anova.scores <- mclapply(1:70,function(each.group){
  select <- which(as.character(ordered.data$bio.phenos$fam_id) %in% train.fam.matrix[each.group,])
  train.dat <- log.bio.counts[select,]
  train.p <- ordered.data$bio.phenos$Volume[select]
  pval.trains <- apply(train.dat,2,function(z) anovaScores(x = z,y = train.p))
  p.adjust(pval.trains,method = "fdr")
},mc.cores = 8)
proc.time() - pt
```

* Now conduct anova on SNPs for each of the 70 groups.

```{r anova.snps}
pt <- proc.time()
snp.anova.scores <- mclapply(1:70,function(each.group){
  select <- which(as.character(ordered.data$bio.phenos$fam_id) %in% train.fam.matrix[each.group,])
  train.dat <- ordered.data$bio.snps[select,]
  train.p <- ordered.data$bio.phenos$Volume[select]
  pval.trains <- apply(train.dat,2,function(z) anovaScores(x = z,y = train.p))
  p.adjust(pval.trains,method = "fdr")
},mc.cores = 8)
proc.time() - pt
```


## Save output

* Create list of objects needed and save

```{r save}

fam.counts <- scale(log2(ordered.data$fam.counts + 1),center=F)

cv.70.data <- list("snp.fdr.anova"=snp.anova.scores,"txpt.fdr.anova"=txpt.anova.scores,
                   "train.index"=train.index.matrix,"test.index"=test.index.matrix,
                   "train.fam.index"=train.fam.matrix,"test.fam.index"=test.fam.matrix,
                   "fam.phenos"=ordered.data$fam.phenos,"all.counts"=fam.counts,
                   "all.snps"=ordered.data$fam.snps)
save(cv.70.data,file = "~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/cv.70.data.RData",compress=T)
```

