---
title: "Visualize LOO predictions"
author: "Adam Festa"
date: "12/18/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

Load starting data and library packages

```{r cars}
library(caret,quietly = T); library(ggpubr,quietly = T)

load("~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/step4_create.70.RData")
load("~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/step7_loo.predictions.RData")
```


## Starting data

* The object cv.70.data contains the train index, count/snp matrices, and phenotypes. The object cv.loo.predictions is a list of predictions for each of the 56 families

* Within each of the 56 lists are predictions for that family using txpts, snps, or txpts + snps for each of the 3 models


```{r pressure, echo=FALSE}
length(cv.loo.predictions)
length(cv.loo.predictions[[1]])
```


## Pearson correlation across all 56 families by Data Type & Model

```{r }

true.breeding.values <- cv.70.data$fam.phenos$Volume
lasso.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$lasso))
apply(lasso.predictions,2,function(x) cor(x,true.breeding.values))

ridge.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$ridge))
apply(ridge.predictions,2,function(x) cor(x,true.breeding.values))

ok.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$ok))
apply(ok.predictions,2,function(x) cor(x,true.breeding.values))

```

## RMSE across all 56 families by Data Type & Model

```{r }
lasso.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$lasso))
apply(lasso.predictions,2,function(x) RMSE(x,true.breeding.values))

ridge.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$ridge))
apply(ridge.predictions,2,function(x) RMSE(x,true.breeding.values))

ok.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$ok))
apply(ok.predictions,2,function(x) RMSE(x,true.breeding.values))
```

## LOO prediction plot

* The best accuracy came from txpts+snps using lasso. Let's plot the results:

```{r}
lasso.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$lasso))
plot(lasso.predictions$txpts.snps,true.breeding.values)
```

* Fitting LOESS

```{r}
f.dat <- data.frame("prediction"=lasso.predictions$txpts.snps,"true.values"=true.breeding.values)
ggplot(f.dat,aes(x=prediction,y=true.values)) + geom_point() + geom_smooth(method="loess") +
    stat_cor() + theme_pubclean() + geom_vline(xintercept = 117.8573) + geom_hline(yintercept = 117.8573) 
```

* Fitting LM

```{r}
f.dat <- data.frame("prediction"=lasso.predictions$txpts.snps,"true.values"=true.breeding.values)
ggplot(f.dat,aes(x=prediction,y=true.values)) + geom_point() + geom_smooth(method="lm") +
    stat_cor() + theme_pubclean() + geom_vline(xintercept = 117.8573) + geom_hline(yintercept = 117.8573) 
```


* The intercept is drawn at the population mean for the true breeding values


## Residual by family

```{r}
lasso.predictions <- do.call(rbind,lapply(cv.loo.predictions,function(x) x$lasso))
plot(lasso.predictions$txpts.snps-true.breeding.values)
```


```{r}
f.dat <- data.frame("fam.residual"=lasso.predictions$txpts.snps - true.breeding.values,"vol.accuracy"=cv.70.data$fam.phenos$`Volume Accuracy`)
ggplot(f.dat,aes(x=fam.residual,y=vol.accuracy)) + geom_point() + geom_smooth(method="auto") +
    stat_cor() + theme_pubclean() 
```
