---
title: "LGEP vs. EW Data"
author: "Adam Festa"
date: "11/27/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

Conduct anova for each SNP/transcript using biological replicates and fdr adjust

## Load data

* Load the object from Step 1 in predictions "ordered.data.RData"

```{r pressure}
library(caret); library(parallel)
load("~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/ordered.data.RData")
names(ordered.data)
```

## Create training set

* Training set will be all LGEP families, including those from EW which overlapped

```{r anova}
ordered.data$bio.counts <- scale(log2(ordered.data$bio.counts + 1),center=F)
rm.ew.rows <- which(as.character(ordered.data$bio.phenos$fam_id) %in% ordered.data$ew.fams)
train.bio.counts <- ordered.data$bio.counts[-rm.ew.rows,]
train.bio.snps <- ordered.data$bio.snps[-rm.ew.rows,]
train.bio.phenos <- ordered.data$bio.phenos[-rm.ew.rows,]
rm(rm.ew.rows)
dim(train.bio.counts);dim(train.bio.snps);dim(train.bio.phenos)
```

## Create test and train for prediction

* Prediction will be done on family-mean values, so create test/train on family mean objects

```{r create.train}
rm.ew.fams <- which(as.character(ordered.data$fam.phenos$fam_id) %in% ordered.data$ew.fams)
ordered.data$fam.counts <- scale(log2(ordered.data$fam.counts + 1),center=F)
train.lgep.counts <- ordered.data$fam.counts[-rm.ew.fams,]
train.lgep.snps <- ordered.data$fam.snps[-rm.ew.fams,]
train.lgep.phenos <- ordered.data$fam.phenos[-rm.ew.fams,]

test.ew.counts <- ordered.data$fam.counts[rm.ew.fams,]
test.ew.snps <- ordered.data$fam.snps[rm.ew.fams,]
test.ew.phenos <- ordered.data$fam.phenos[rm.ew.fams,]
```

## Save output for prediction across batch

* Create list of all objects needed for prediction of EW using LGEP

```{r save}
lgep.vs.ew.data <- list("train.counts"=train.lgep.counts,"train.snps"=train.lgep.snps,"train.phenos"=train.lgep.phenos,"test.counts"=test.ew.counts,"test.snps"=test.ew.snps,"test.phenos"=test.ew.phenos)

save(lgep.vs.ew.data,file="~/Desktop/Alt Approaches/Breeding-Value-Prediction/shared/lgep.vs.ew.data.RData",compress=T)
```
