---
title: "Data Restructure"
author: "Adam Festa"
date: "11/27/2018"
output: 
  html_document: 
    theme: flatly
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

The purpose of this script is to take the data that was produced by normalization & SNP filtering to match to the phenotypes so that all are in same order.  Additionally, the crossover families are identified and family mean data sets are generated.

* Script was ran on local computer, but all files are on titan

## Load SNPs, Transcripts & Phenos

* Loads 192x46407 SNPs identified by snpEff and specify row/column names

```{r load.snps}
library(parallel)
snps <- read.table("../../../shared/192.biorep.Q30.snps.no.miss.maf05.ann.high.mod.012",row.names = 1)
rownames(snps) <- read.table("../../../shared/192.biorep.Q30.snps.no.miss.maf05.ann.high.mod.012.indv")[,1]
colnames(snps) <- paste0(read.table("../../../shared/192.biorep.Q30.snps.no.miss.maf05.ann.high.mod.012.pos")[,1],
                         "_",
                         read.table("../../../shared/192.biorep.Q30.snps.no.miss.maf05.ann.high.mod.012.pos")[,2])
```

* Fix the SNP rownames:

```{r fix.snp.names}
head(rownames(snps))
this <- gsub("[^0-9\\.]", "", rownames(snps))
this <- gsub("\\.","",this)
this <- as.numeric(this)
this[1:48] <- this[1:48] + 1000
rownames(snps) <- this ; rm(this)
head(rownames(snps))
```

* Loads trancript normalization returned from asreml

```{r load.txpts}
load("../../../shared/Step3_LMM_animal.RData")
```

* The rownames need to be fixed:

```{r fix.txpt.names}
head(rownames(asreml.counts))
rownames(asreml.counts) <- gsub(pattern = "ped\\(animal_id, var = T)_",replacement = "",x = rownames(asreml.counts))
head(rownames(asreml.counts))
asreml.counts <- asreml.counts[-c(1:94),]
```

* Load all phenos (tech reps)

```{r load.phenos}
load("../../../disk6directory/resources/expt.dat.720.RData")
```

## Consolidate phenos to biological reps

* Reduce the tech reps phenos to only include bio rep entries found in SNPs/transcripts

```{r reduce.phenos}
# Find column names
#colnames(expt.dat.720)
# unique set of animal_id, fam_id, Volume, & Height
bio.phenos <- unique(expt.dat.720[,c(1,2,13,17,18)])

# subset phenos to only include entries which passed transcript qc
bio.phenos <- bio.phenos[which(as.character(bio.phenos$animal_id) %in% rownames(asreml.counts)),]

# remove missing phenotypes
bio.phenos <- bio.phenos[-c(which(is.na(bio.phenos$Volume) == T)),]

# Reduce snps to only include those with phenotypes 
snps <- snps[which(rownames(snps) %in% as.character(bio.phenos$animal_id)),]
dim(snps)
# 189 x 46047

# Reduce phenos to only include those within snps
bio.phenos <- bio.phenos[which(as.character(bio.phenos$animal_id) %in% rownames(snps)),]
dim(bio.phenos)
#189x4

# Reduce transcripts to include those within phenos
asreml.counts <- asreml.counts[which(rownames(asreml.counts) %in% bio.phenos$animal_id),]
dim(asreml.counts)
#189x70109
```

## Match all data sets

* Identify the order of the bio.phenos object and match bio counts and snps to that order

```{r match.order}
# Match txpt rownames to bio.phenos
this.order <- match(as.character(bio.phenos$animal_id),rownames(asreml.counts))
identical(rownames(asreml.counts)[this.order],as.character(bio.phenos$animal_id))
#TRUE
asreml.counts <- asreml.counts[this.order,]

# Match snps rownames to bio.phenos
this.order <- match(as.character(bio.phenos$animal_id),rownames(snps))
identical(rownames(snps)[this.order],as.character(bio.phenos$animal_id))
#TRUE
snps <- snps[this.order,]
rm(this.order)
```

## Obtain family level data

* Generate family-mean level phenotypes, transcript counts, & SNPs

```{r fam.means}
# Fam phenos will be the unique of fam_id, Volume, Height
fam.phenos <- unique(bio.phenos[,-1])
rownames(fam.phenos) <- fam.phenos$fam_id

# Consolidate bio transcripts
fam.counts <- mclapply(1:56,function(each.fam){
  these.rows <- which(as.character(bio.phenos$fam_id) %in% rownames(fam.phenos)[each.fam])
  if(length(these.rows) > 1){
  apply(asreml.counts[these.rows,],2,mean)} else {
    asreml.counts[these.rows,]
  }
},mc.cores=8)
fam.counts <- do.call(rbind,fam.counts)
rownames(fam.counts) <- fam.phenos$fam_id
dim(fam.counts)
#56 70109

# Consolidate bio SNPs
fam.snps <- mclapply(1:56,function(each.fam){
  these.rows <- which(as.character(bio.phenos$fam_id) %in% rownames(fam.phenos)[each.fam])
  if(length(these.rows) > 1){
  apply(snps[these.rows,],2,mean)} else {
    snps[these.rows,]
  }
},mc.cores=8)
fam.snps <- do.call(rbind,fam.snps)
rownames(fam.snps) <- fam.phenos$fam_id
dim(fam.snps)
#56 46407
```

## Identify crossover families between LGEP & EW

* Find the families which were grown in both batches
```{r crossover.fams}
crossover.fam.table <- unique(expt.dat.720[,c(2,5,17)])
crossover.fam.table <- crossover.fam.table[which(as.character(crossover.fam.table$fam_id) %in% rownames(fam.phenos)),]
crossover.fams <- table(as.character(crossover.fam.table$fam_id))
crossover.fams <- names(crossover.fams)[which(crossover.fams == 2)]
crossover.fams
```

* Find the EW families

```{r ew.fams}
ew.fams <- as.character(crossover.fam.table$fam_id[which(as.character(crossover.fam.table$batch) %in% "EW")])
ew.fams <- ew.fams[-which(ew.fams %in% crossover.fams)]
rm(crossover.fam.table)
ew.fams
lgep.fams <- as.character(fam.phenos$fam_id[-which(fam.phenos$fam_id %in% ew.fams)])
```

## Save all data in shared directory

* Create a list of all data objects and save

```{r save}
ordered.data <- list("bio.counts"=asreml.counts,"bio.snps"=snps,"bio.phenos"=bio.phenos,"fam.counts"=fam.counts,"fam.snps"=fam.snps,"fam.phenos"=fam.phenos,
     "ew.fams"=ew.fams,"lgep.fams"=lgep.fams,"crossover.fams"=crossover.fams)

save(ordered.data,file = "../../../shared/ordered.data.RData",compress = T)
```



