---
title: "LGEP vs. EW Prediction"
author: "Adam Festa"
date: "2/4/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
## Purpose / load starting data
  
The purpose of this script is to conduct predictions with ridge, lasso, and OmicKriging on the EW families using the LGEP families
 
Load the LGEP vs EW data & R packages
 
```{r load.data, message=FALSE, warning=FALSE, include=FALSE}
suppressPackageStartupMessages(expr = c(library(knitr),library(glmnet,quietly = T),library(ggplot2,quietly = T,verbose = F),
library(doParallel), library(ggplot2),library(ggpubr), library(OmicKriging), library(caret)))
load("../../../shared/lgep.vs.ew.ridge.lasso.ok.V3.RData")
```

## Results

### Table of results

A table of results including the pearson correlation, spearman correlation, and RMSE using transcripts, SNPs, or transcripts + SNPs to predict the EW families

```{r,echo=F}
lasso.cor = c(cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$lasso.pred),
cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$lasso.pred),
cor(ew.lgep.results$snps$true.phenos,(ew.lgep.results$txpts$lasso.pred+ew.lgep.results$snps$lasso.pred)/2))

ridge.cor = c(cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$ridge.pred),
cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$ridge.pred),
cor(ew.lgep.results$snps$true.phenos,c(ew.lgep.results$txpts$ridge.pred + ew.lgep.results$snps$ridge.pred)/2))

lasso.spear = c(cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$lasso.pred,method="spearman"),
cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$lasso.pred,method="spearman"),
cor(ew.lgep.results$snps$true.phenos,c(ew.lgep.results$txpts$lasso.pred + ew.lgep.results$snps$lasso.pred)/2,method="spearman"))

ridge.spear = c(cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$ridge.pred,method="spearman"),
cor(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$ridge.pred,method="spearman"),
cor(ew.lgep.results$snps$true.phenos,c(ew.lgep.results$txpts$ridge.pred + ew.lgep.results$snps$ridge.pred)/2,method="spearman"))

lasso.rmse = c(RMSE(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$lasso.pred),
RMSE(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$lasso.pred),
RMSE(ew.lgep.results$snps$true.phenos,c(ew.lgep.results$txpts$lasso.pred + ew.lgep.results$snps$lasso.pred)/2))

ridge.rmse = c(RMSE(ew.lgep.results$snps$true.phenos,ew.lgep.results$snps$ridge.pred),
RMSE(ew.lgep.results$snps$true.phenos,ew.lgep.results$txpts$ridge.pred),
RMSE(ew.lgep.results$snps$true.phenos,c(ew.lgep.results$txpts$ridge.pred + ew.lgep.results$snps$ridge.pred)/2))

dat.frame <- cbind(lasso.cor,lasso.spear,lasso.rmse,ridge.cor,ridge.spear,ridge.rmse)
rownames(dat.frame) <- c("snps","txpts","txpts+snps")
kable(dat.frame)
```

### Plot top results

```{r}
pred.dat <- data.frame("prediction"=(ew.lgep.results$snps$lasso.pred + ew.lgep.results$txpts$lasso.pred)/2 ,"truth"=ew.lgep.results$snps$true.phenos)
ggplot(pred.dat,aes(x=X1,y=truth)) + geom_point() + theme_pubclean() + geom_smooth(method = "lm") + theme_pubclean() + stat_cor() + ylab("True BV") + xlab("Predicted BV")
```

### Plot transcripts alone

```{r}
pred.dat <- data.frame("prediction"=ew.lgep.results$txpts$lasso.pred,"truth"=ew.lgep.results$snps$true.phenos)
ggplot(pred.dat,aes(x=X1,y=truth)) + geom_point() + theme_pubclean() + geom_smooth(method = "lm") + theme_pubclean() + stat_cor() + ylab("True BV") + xlab("Predicted BV")
```

### Lasso feature variables

Number of variables selected for SNPs and transcripts

```{r}
length(ew.lgep.results$snps$lasso.feats)
length(ew.lgep.results$txpts$lasso.feats)
```

* There are 59 SNPs and 60 transcripts which were used for prediction

Remove SNP position from name to get transcript name that snp was found

```{r}
# First remove the position of snps within transcripts from SNP lasso features
snp.features.no.pos <- sub("_[^_]+$", "", ew.lgep.results$snps$lasso.feats)
length(unique(snp.features.no.pos))
#table(snp.features.no.pos)
```
 
* There are 58 unique transcripts in which a SNP was found, the transcript which had two snps in it was est_Pita_34350202

Check for any transcripts which were selected using both transcripts & SNPs
```{r}
which(ew.lgep.results$txpts$lasso.feats %in% unique(snp.features.no.pos))
```

No transcripts were selected in both snps & transcripts
