---
title: 'CARET: Predict EW Volume BV’s with LGEP: SNPs 012'
author: "Adam Festa"
date: "6/10/2018"
output: 
  html_document: 
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(caret));suppressPackageStartupMessages(library(caretEnsemble)); library(parallel); library(Metrics)
setwd("/mnt/mnt/media/disk6/ARF/shared")
```

## Load data

* Load the --012 output produced during the SNP filtering.

* File loaded below has no missing snps, MAF > .05, and are of at least Q30.

* Additionally, we get the rownames to match the names that are present in the phenos object

```{r load.data}
# Load matrix, indv file, and snp position file (indv file rownames & snp_pos colnames)
out.012 <- read.table("./snps/012/Q30.snps.nomiss.maf05.012",header = F)
rownames(out.012) <- as.character(read.table("./snps/012/Q30.snps.nomiss.maf05.012.indv")[,1])
colnames(out.012) <- paste0(as.character(read.table("./snps/012/Q30.snps.nomiss.maf05.012.pos")[,1]),
                            "_",
                            read.table("./snps/012/Q30.snps.nomiss.maf05.012.pos")[,2])

# Remove any extra "_" characters in the sample names
rel.colnames <- gsub(rownames(out.012),pattern = "_",replacement = "")
# Identify those matching the EW samples and replace the following statement
rel.colnames.ew <- gsub(rel.colnames,pattern = ".ew.bio.rep.merge.bam",replacement = "")

# We need to add "1000" to these names in order to match the phenos, 
# so here just subset the ones which when the string is forced to be numeric, it has numbers
ew.bio.reps <- which(!is.na(as.numeric(rel.colnames.ew)))

# The remaining 192 samples are LGEP
lgep.bio.reps <- c(1:192)[-ew.bio.reps]
# Now add 1000 to the EW sample names
mc.ew <- as.numeric(rel.colnames.ew[ew.bio.reps]) + 1000


# Subset the lgep bio reps, replace the matching pattern, and remove the index character attached to the sample number
mc.lgep <- rel.colnames[lgep.bio.reps]
mc.lgep <- gsub(pattern = "*.lgep.bio.rep.merge.bam",replacement = "",x = mc.lgep)
mc.lgep <- substr(mc.lgep,1,nchar(mc.lgep)-1)

# Finally replace the original column names with the adjusted colnames and remove vars
rel.colnames[ew.bio.reps] <- as.character(mc.ew)
rel.colnames[lgep.bio.reps] <- mc.lgep
head(rel.colnames)
rownames(out.012) <- rel.colnames
rm(rel.colnames,rel.colnames.ew,ew.bio.reps,lgep.bio.reps,mc.ew,mc.lgep)

## Load phenotype data ####
#load("~/Documents/Grad_Projects/Breeding-Value-Prediction/disk6directory/resources/expt.dat.720.RData")
load("/repos/Breeding-Value-Prediction/disk6directory/resources/expt.dat.720.RData")
expt.dat.720$animal_id <- as.character(expt.dat.720$animal_id)
expt.dat.720$fam_id <- as.character(expt.dat.720$fam_id)
expt.dat.720$batch <- as.character(expt.dat.720$batch)

# Create bio phenos from individuals with no missing Volume and who have unique, fam*animal*batch*vol combo
bio.phenos <- unique(expt.dat.720[which(!is.na(expt.dat.720$Volume) ==T),c("fam_id","animal_id","batch","Volume")])
```

* The bio phenos animal id and 012 matrix have the same style names but are not in the same order. Match 012 genos to phenotype data

```{r match}
# Subset bio.phenos to only include individuals within the snp mat
keep.index <-  which(rownames(out.012) %in% bio.phenos$animal_id)
out.012.subset <- out.012[keep.index,]

bio.pheno.subset <- bio.phenos[which(bio.phenos$animal_id %in% rownames(out.012.subset)),]
identical(bio.pheno.subset$animal_id[match(rownames(out.012.subset),bio.pheno.subset$animal_id)],rownames(out.012.subset))

bio.pheno.subset <- bio.pheno.subset[match(rownames(out.012.subset),bio.pheno.subset$animal_id),]
rownames(bio.pheno.subset) <- bio.pheno.subset$animal_id
rm(bio.phenos,out.012,keep.index)
```


## Generate family-level data

* Generate family phenos

```{r fam.phenos}
fam.phenos <- unique(bio.pheno.subset[,c("fam_id","Volume")])
rownames(fam.phenos) <- fam.phenos$fam_id
u.fams <- unique(fam.phenos$fam_id)
```

* Generate family mean SNP matrix using the FULL bio 012 matrix

```{r fam.mean}
fam.012 <- mclapply(1:nrow(fam.phenos),function(each.fam){
  these.rows <- which(bio.pheno.subset$fam_id %in% u.fams[each.fam])
    if(length(these.rows) < 2) { unlist(out.012.subset[these.rows,])
      } else {
         apply(out.012.subset[these.rows,],2,mean)
    }
},mc.cores=30)
fam.012 <- do.call(rbind,fam.012)
```


## Create test/train matricies

* Some of the filtering methods will use phenotypes, while others will not, but we can go ahead and split the LGEP and EW into train/test for those filtering methods that do use phenos

```{r test.train}
# Identify crossover families between batches
crossover.fams <- unique(bio.pheno.subset[,c("fam_id","Volume","batch")])
crossover.fams <- names(which(table(crossover.fams$fam_id) ==2))

# Subset the unique list of families that are in the LGEP or EW
ew.fams <- unique(bio.pheno.subset$fam_id[which(bio.pheno.subset$batch == "EW")])
lgep.fams <- unique(bio.pheno.subset$fam_id[which(bio.pheno.subset$batch == "LGEP")])

# Remove LGEP families from the EW families (i.e. remove crossover families)
ew.fams <- ew.fams[-which(ew.fams %in% crossover.fams)]

# Create a training set from the biological replicates using the lgep and ew families
train.p <- bio.pheno.subset$Volume[which(bio.pheno.subset$fam_id %in% lgep.fams)]
train.dat <- out.012.subset[which(bio.pheno.subset$fam_id %in% lgep.fams),]
colnames(train.dat) <- 1:ncol(train.dat)
```

* The object `train.dat` now contains the LGEP 012 output and the `out.012.subset` object contains the full 189 biological replicates.

## Apply filters

* Apply caret filters including:

    * ANOVA scores: These are generated using phenotypes, so only the LGEP will be used to identify a subset to predict
    
    * BCV:  This may be generated without phenotypes, so the whole dataset will be used.
    
    * nearzerovar: This may also be used without phenotypes, so the whole dataset will be used.

### Generate anova scores

* Use caret package to generate anova scores with the LGEP 012 counts and phenos (i.e. `train.dat` & `train.phenos` objects)

```{r lgep.anova}
# Using the biological replicate training set, apply anovaScores to each SNP
anova_scores <- unlist(mclapply(1:ncol(train.dat), function(each.col){ 
  anovaScores(x = train.dat[,each.col],y = train.p)},mc.cores=30))

# Generate a histogram of these anova scores
hist(anova_scores, main="Unadjusted ANOVA scores using LGEP 012 and phenos",xlab ="anova scores",ylab="# of SNPs")
# How many of anova scores are less than .05 (unadjusted)
length(which(anova_scores < .05))
# How many of anova scores are less than .05 (adjusted)
length(which(p.adjust(anova_scores) < .05))
```

* A total of 23,334 SNPs are < .05 with unadjusted p-values and approximately 88 SNPs are found with FDR adjusted p-values < .05

### Filter BCV

* Here we will use the whole 012 data set to identify SNPs who have high BCV relative to all samples (BCV= sd/mean)

```{r bcv.stats}
# Estimate the variance, sd, mean, and bcv for each SNP
var.snps <- apply(out.012.subset,2,var)
sd.snps <- apply(out.012.subset,2,sd)
mean.snps <- apply(out.012.subset,2,mean)
bcv.snps <- sd.snps/mean.snps

# Take a look at the distribution of bcv values
hist(bcv.snps,main = "Histogram of BCV values for the whole 012 dataset",xlab="BCV value",ylab="# of SNPs")
length(which(bcv.snps > 3))
# Look at the histogram of variance using only those SNPs with bcv > 3
hist(var.snps[which(bcv.snps > 3)],main="Histogram of variance for SNPs with BCV > 3",xlab = "variance estimate",ylab = "# of SNPs")
length(which(var.snps[which(bcv.snps > 3)] < .2))
```

* A total of 9,219 SNPs have BCV values greater than 3.

* There is an interesting set of 624 SNPs which have a BCV greater than 3 but variance less than 0.2


### Filter the nearzerovar NOT RUN

* The 'nearzerovar' function will check for things which have less than 10 or less unique values and 95/5 most common/least common (19:1 ratio).  This subset was interesting in the last

```{r check.vars}
# This near zero var will check for things which have less than 10 or less unique values and 95/5 most common/least common (19:1 ratio)
check.vars <- nearZeroVar(x = out.012.subset)

# How many did not meet these thresholds?
length(check.vars)

## 6,624 SNPs did not meet threshold
```

* A total of 6624 SNPs did not meet this threshold, let's take a look at their BCV values

```{r nearzero.stats}
hist(bcv.snps[check.vars])
```

* There is an odd set which have > 3 BCV. Let's subset those out too.

```{r nearzero.bcv}
near.zero.bcv3.snps <- check.vars[which(bcv.snps[check.vars] > 3)]
length(near.zero.bcv3.snps)
```

* A total of 253 SNPs were identified as being removed by the nearzerovar BUT they had BCV > 3. We will keep those for further inspection.

* Do these 253 SNPs show up as significant in the p-adjusted anova scores on the full data?

```{r near.zero.anova}
summary(p.adjust(anova_scores)[near.zero.bcv3.snps])
```

* No, none of these 253 were identified as being associated with the LGEP

* How do the variance of these SNPs look?

```{r near.zero.var}
summary(var.snps[near.zero.bcv3.snps])
```

* The SNP variance's are all pretty small. We will still keep this set and use during prediction.

## Caret pt1: Pressure test multiple models

* We have 3 various filters to work with including:

    * Anova scores generated using ONLY LGEP
    
    * BCV values generated using ALL data
    
    * Nearzerovar subset identified with ALL data (BCV > 3)

* In this section, we will run all *faster* models on the training set (LGEP families) and look at which models tend to perform best in terms of `R2` & `rmse`

### Run all *faster* models on LGEP training

* Establish the `trainControl` object
```{r t.control}
tC <- trainControl(method="boot632",allowParallel = T, savePredictions="all",verboseIter = TRUE,search="grid")
```

* Generate a list of models to test on LGEP

```{r models}
{algorithmList <- c("bagEarth", "bagEarthGCV", 
"bayesglm", "blackboost", "brnn", "BstLm" , 
"bstTree", "cforest", "ctree", "ctree2", "cubist", "DENFIS", 
"dnn", "earth", "elm", "enet",   "evtree", 
"extraTrees",  "gamLoess",  "gaussprLinear", "gaussprPoly", "gaussprRadial", 
"gcvEarth","glm", "glmboost", "glmnet", "icr", "kernelpls", 
"kknn", "knn",  "krlsRadial", "lars" , "lasso", 
"leapBackward", "leapForward", "leapSeq", "lm", "M5", "M5Rules", 
"mlpWeightDecay", "neuralnet" , "partDSA", 
"pcaNNet", "pcr", "penalized", "pls", "plsRglm", "ppr", 
"qrf" , "ranger",  "rf", "rfRules", "rbfDDA",
"ridge", "rlm", "rpart", "rpart2", "rqlasso", 
"rqnc", "RRF", "RRFglobal",  "rvmPoly", "rvmRadial", 
"SBC", "simpls", "spls", "superpc" , 
"svmLinear", "svmLinear2", "svmPoly", "svmRadial", "svmRadialCost", 
"treebag", "widekernelpls", "WM", "xgbLinear", 
"xgbTree", "xyf")}


algorithmList <- c("glm","glmnet", "lm")
```

* Execute the models

```{r}
LGEP.train.dat <- fam.012[which(fam.phenos$fam_id %in% lgep.fams),which(p.adjust(anova_scores) < .05)]
colnames(LGEP.train.dat) <- colnames(fam.012[,which(p.adjust(anova_scores) < .05)])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
```

```{r execute.models}
library(doParallel); cl <- makeCluster(detectCores()); registerDoParallel(cl)

t2 <- lapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = trainControl(method = "boot632"))
	}
)


r2 <- lapply(1:length(t2), function(i) 
		{cat(sprintf("%-20s",(algorithmList[i])));
		cat(round(t2[[i]]$results$Rsquared[which.min(t2[[i]]$results$RMSE)],4),"\t");
		cat(round(t2[[i]]$results$RMSE[which.min(t2[[i]]$results$RMSE)],4),"\t")
		cat(t2[[i]]$times$everything[3],"\n")
		}
)
stopCluster(cl); registerDoSEQ();
```

```
#trainControl <- trainControl(method="LOOCV",allowParallel = T, savePredictions="all",verboseIter = TRUE,search="grid",number=5)
models <- caretList(y=train.p,x=train.dat,  trControl=trainControl, methodList=algorithmList,continue_on_fail = T)
results <- resamples(models)
#save(results,file="./caret_1cv_alldat.ew.lgep.Rdata")
summary(results)
# Box plots to compare models
scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(results, scales=scales)
* Identify top R^2 outputs
out <- summary(results)
out <- out$statistics$RMSE
```
```

* Extract the results of the models

```{r xtract.sum.mods}
# preallocate data types
i = 1; MAX = length(t2);
x1 <- character() # Name
x2 <- numeric()   # R2
x3 <- numeric()   # RMSE
x4 <- numeric()   # time [s]
x5 <- character() # long model name
 
# fill data and check indexes and NA
for (i in 1:length(t2)) {
    x1[i] <- t2[[i]]$method
    x2[i] <- as.numeric(t2[[i]]$results$Rsquared[which.min(t2[[i]]$results$RMSE)])
    x3[i] <- as.numeric(t2[[i]]$results$RMSE[which.min(t2[[i]]$results$RMSE)])
    x4[i] <- as.numeric(t2[[i]]$times$everything[3])
    x5[i] <- t2[[i]]$modelInfo$label
}
  
# coerce to data frame
df1 <- data.frame(x1,x2,x3,x4,x5, stringsAsFactors=FALSE)

# call web browser output with sortable column names
datatable(df1,  options = list(
		columnDefs = list(list(className = 'dt-left', targets = c(0,1,2,3,4,5))),
		pageLength = MAX,
  		order = list(list(2, 'desc'))),
		colnames = c('Num', 'Name', 'R^2', 'RMSE', 'time [s]', 'Model name'),
	        caption = paste('Regression results from caret models',Sys.time()),
	        class = 'cell-border stripe')  %>% 	       
	        formatRound('x2', 3) %>%  
	        formatRound('x3', 3) %>%
	        formatRound('x4', 3) %>%
		    formatStyle(2,
		    background = styleColorBar(x2, 'steelblue'),
		    backgroundSize = '100% 90%',
		    backgroundRepeat = 'no-repeat',
		    backgroundPosition = 'center'
)

```



# Stacking Algorithms - Run multiple algos in one call.((


# Create groups for training
#grp.list <- vector("list")
#new.l <- 1:56
#for(each in 1:7){
#grp.list[[each]] <- sample(new.l,replace = F,size = 8)
#new.l <- new.l[-unlist(grp.list[[each]])]
#}

# Set up the train control
trainControl <- trainControl(method="repeatedcv",repeats=50,allowParallel = T, savePredictions="all",verboseIter = TRUE,search="grid")
#algorithmList <- c('earth','bagEarth','glmnet','xgbDART','xgbLinear','glmboost','treebag','rf','BstLm')
#algorithmList <- c('bayesglm','glm')'svmLinear2'superpc'rrlda, rlm ,foba,ridge,RRF,rfRules,glm.nb

algorithmList <- c('glmnet','glm')

models <- caretList(y=train.p,x=train.dat,  trControl=trainControl, methodList=algorithmList,continue_on_fail = T)
results <- resamples(models)
summary(results)

# Box plots to compare models
scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(results, scales=scales)

trainControl <- trainControl(method="LOOCV",allowParallel = T, savePredictions="all",verboseIter = TRUE,search="grid")
pred <- vector()
  set.seed(100)
  train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
  train.dat <- fam.012.cormat[which(fam.phenos$fam_id %in% lgep.fams),]
  #train.p <- fam.phenos$Volume[-check.samp]
  #train.dat <- fam.012.cormat[-check.samp,]
  colnames(train.dat) <- 1:ncol(train.dat)
  #mod1 <- train(y=train.p,x = train.dat,trControl=trainControl,method="ridge")
  mod1 <- train(y=train.p,x = train.dat,trControl=trainControl,method="glmnet")
  #mod1 <- train(y=train.p,x = train.dat,trControl=trainControl,method="BstLm")
  #mod1
  test.dat <- fam.012.cormat; colnames(test.dat) <- colnames(train.dat)
  #pred <- c(pred,predict(mod1,test.dat)[check.samp])
  pred <- c(pred,predict(mod1,test.dat)[which(fam.phenos$fam_id %in% ew.fams)])
  #plot(pred,fam.phenos$Volume[1:check.samp])
  plot(pred,fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)])
}
abline(h=mean(fam.phenos$Volume),v=mean(pred))
plot(pred[-c(2,15,32,48)],fam.phenos$Volume[-c(2,15,32,48)])
