---
title: 'CARET: Predict EW Volume BV’s with LGEP: SNPs 012'
author: "Adam Festa"
date: "6/10/2018"
output: 
  html_document: 
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}

plot.mean.model.preds <- function(model.list=t2){
results.df <- do.call(cbind,lapply(t2,function(i){(predict(i,EW.test.dat))}))
colnames(results.df) <- algorithmList;rownames(results.df) <- ew.fams
results.df<- data.frame(cbind("fam"=EW.test.p,results.df))

library(tidyr)
data_long <- gather(data = results.df, model, prediction, "glmnet":"svmRadialCost", factor_key=F)

tgc <- summarySE(data_long, measurevar="prediction",groupvars=c("fam"))
pd <- position_dodge(0.1)
ggplot(tgc, aes(x=fam, y=prediction, colour=fam, group=fam)) + 
  geom_errorbar(aes(ymin=prediction-se, ymax=prediction+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=3, shape=21, fill="white") + # 21 is filled circle
  xlab("True value") +
  ylab("Predicted Value") +
  ggtitle("Average Prediction across 12 models") +
  expand_limits(y=100) +                        # Expand y range
  scale_y_continuous() +         # Set tick every 4
  theme_bw() +
  theme(legend.justification=c(1,0),
        legend.position=c(1,0))               # Position legend in bottom right 
}
create.data.table <- function(model.list=t2) {
  
# preallocate data types
i = 1; MAX = length(model.list);
x1 <- character() # Name
x2 <- numeric()   # R2
x3 <- numeric()   # RMSE
x4 <- numeric()   # time [s]
x5 <- character() # long model name

# fill data and check indexes and NA
for (i in 1:length(model.list)) {
  x1[i] <- model.list[[i]]$method
  x2[i] <- as.numeric(model.list[[i]]$results$Rsquared[which.min(model.list[[i]]$results$Rsquared)])
  x3[i] <- as.numeric(model.list[[i]]$results$RMSE[which.min(model.list[[i]]$results$RMSE)])
  x4[i] <- as.numeric(model.list[[i]]$times$everything[3])
  x5[i] <- model.list[[i]]$modelInfo$label
}

# coerce to data frame
df1 <- data.frame(x1,x2,x3,x4,x5, stringsAsFactors=FALSE)

# call web browser output with sortable column names
datatable(df1,  options = list(
  columnDefs = list(list(className = 'dt-left', targets = c(0,1,2,3,4,5))),
  pageLength = MAX,
  order = list(list(2, 'desc'))),
  colnames = c('Num', 'Name', 'R^2', 'RMSE', 'time [s]', 'Model name'),
  caption = paste('Regression results from caret models',Sys.time()),
  class = 'cell-border stripe')  %>% 	       
  formatRound('x2', 3) %>%  
  formatRound('x3', 3) %>%
  formatRound('x4', 3) %>%
  formatStyle(2,
              background = styleColorBar(x2, 'steelblue'),
              backgroundSize = '100% 90%',
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center')
}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(caret));suppressPackageStartupMessages(library(caretEnsemble)); library(parallel); library(Metrics); library(DT)
setwd("/mnt/media/disk6/ARF/shared")
#setwd("~/Documents/Grad_Projects/BV_Prediction/mnt/media/disk6/ARF/shared")
```

## Load data

* First, load the --012 output produced during the SNP filtering.

* File loaded below has no missing snps, MAF > .05, and are of at least Q30.

* Additionally, the data frame contains rownames for bioloigcal samples and colnames as snp name/position

```{r load.snp.dat}
load("/mnt/bio.012.df.RData")
```

* Now load the phenotypes

```{r load phenos}
## Load phenotype data ####
#load("~/Documents/Grad_Projects/Breeding-Value-Prediction/disk6directory/resources/expt.dat.720.RData")
load("/mnt/Breeding-Value-Prediction/disk6directory/resources/expt.dat.720.RData")
expt.dat.720$animal_id <- as.character(expt.dat.720$animal_id)
expt.dat.720$fam_id <- as.character(expt.dat.720$fam_id)
expt.dat.720$batch <- as.character(expt.dat.720$batch)

# Create bio phenos from individuals with no missing Volume and who have unique, fam*animal*batch*vol combo
bio.phenos <- unique(expt.dat.720[which(!is.na(expt.dat.720$Volume) ==T),c("fam_id","animal_id","batch","Volume")])
```

* The bio phenos animal id and 012 matrix have the same style names but are not in the same order. Match 012 genos to phenotype data

```{r match}
# Index the snp data frame to only include individuals which have phenotypes
keep.index <-  which(rownames(bio.012.df) %in% bio.phenos$animal_id)

# Keep only those rows
bio.012.df <- bio.012.df[keep.index,]

bio.phenos <- bio.phenos[which(bio.phenos$animal_id %in% rownames(bio.012.df)),]
identical(bio.phenos$animal_id[match(rownames(bio.012.df),bio.phenos$animal_id)],rownames(bio.012.df))

bio.phenos <- bio.phenos[match(rownames(bio.012.df),bio.phenos$animal_id),]
rownames(bio.phenos) <- bio.phenos$animal_id
rm(keep.index)
```


## Generate family-level data

* Generate family phenos

```{r fam.phenos}
fam.phenos <- unique(bio.phenos[,c("fam_id","Volume")])
rownames(fam.phenos) <- fam.phenos$fam_id
u.fams <- unique(fam.phenos$fam_id)
```

* The family mean SNP matrix was generated once using the FULL bio 012 matrix and then saved. It's loaded here

```{r load.fam}
load("/mnt/fam.012.df.RData")
```

## Create test/train matricies

* Some of the filtering methods will use phenotypes, while others will not, but we can go ahead and split the LGEP and EW into train/test for those filtering methods that do use phenos

```{r}
# Identify crossover families between batches
crossover.fams <- unique(bio.phenos[,c("fam_id","Volume","batch")])
crossover.fams <- names(which(table(crossover.fams$fam_id) ==2))

# Subset the unique list of families that are in the LGEP or EW
ew.fams <- unique(bio.phenos$fam_id[which(bio.phenos$batch == "EW")])
lgep.fams <- unique(bio.phenos$fam_id[which(bio.phenos$batch == "LGEP")])

# Remove LGEP families from the EW families (i.e. remove crossover families)
ew.fams <- ew.fams[-which(ew.fams %in% crossover.fams)]

# Create a training set from the biological replicates using the lgep and ew families
train.p.bio <- bio.phenos$Volume[which(bio.phenos$fam_id %in% lgep.fams)]
train.dat.bio <- bio.012.df[which(bio.phenos$fam_id %in% lgep.fams),]
#colnames(train_d.bio) <- 1:ncol(train.dat)
```

* The object `train.dat.bio` now contains the LGEP 012 output and the `bio.012.df` object contains the full 189 biological replicates.

## Apply filters

* Apply caret filters including:

    * ANOVA scores: These are generated using phenotypes, so only the LGEP will be used to identify a subset to predict
    
    * BCV:  This may be generated without phenotypes, so the whole dataset will be used.
    
    * nearzerovar: This may also be used without phenotypes, so the whole dataset will be used.

### Generate anova scores

* Use caret package to generate anova scores with the LGEP 012 counts and phenos (i.e. `train.dat` & `train.phenos` objects)

```
# Using the biological replicate training set, apply anovaScores to each SNP
LGEP.anova.scores <- unlist(mclapply(1:ncol(train.dat.bio), function(each.col){ 
  anovaScores(x = train.dat.bio[,each.col],y = train.p.bio)},mc.cores=32))
save(LGEP.anova.scores,file="/mnt/LGEP.anova.scores.RData",compress = T)
```

```{r lgep.anova}
load("/mnt/LGEP.anova.scores.RData")
# Generate a histogram of these anova scores
hist(LGEP.anova.scores, main="Unadjusted ANOVA scores using LGEP 012 and phenos",xlab ="anova scores",ylab="# of SNPs")
# How many of anova scores are less than .05 (unadjusted)
length(which(LGEP.anova.scores < .05))
# How many of anova scores are less than .05 (adjusted)
length(which(p.adjust(LGEP.anova.scores) < .05))
```

* A total of 23,334 SNPs are < .05 with unadjusted p-values and approximately 88 SNPs are found with FDR adjusted p-values < .05

### Filter BCV

* Here we will use the whole 012 data set to identify SNPs who have high BCV relative to all samples (BCV= sd/mean)

```{r bcv.stats}
# Estimate the variance, sd, mean, and bcv for each SNP
var.snps <- apply(bio.012.df,2,var)
sd.snps <- apply(bio.012.df,2,sd)
mean.snps <- apply(bio.012.df,2,mean)
bcv.snps <- sd.snps/mean.snps

# Take a look at the distribution of bcv values
hist(bcv.snps,main = "Histogram of BCV values for the whole 012 dataset",xlab="BCV value",ylab="# of SNPs")
length(which(bcv.snps > 3))
# Look at the histogram of variance using only those SNPs with bcv > 3
hist(var.snps[which(bcv.snps > 3)],main="Histogram of variance for SNPs with BCV > 3",xlab = "variance estimate",ylab = "# of SNPs")
length(which(var.snps[which(bcv.snps > 3)] < .2))
interesting.bcv.snps <- which(var.snps[which(bcv.snps > 3)] < .2)
```

* A total of 9,219 SNPs have BCV values greater than 3.

* There is an interesting set of 624 SNPs which have a BCV greater than 3 but variance less than 0.2


### Filter the nearzerovar

* The 'nearzerovar' function will check for things which have less than 10 or less unique values and 95/5 most common/least common (19:1 ratio).  This subset was interesting in the last

```{r check.vars}
# This near zero var will check for things which have less than 10 or less unique values and 95/5 most common/least common (19:1 ratio)
check.vars <- nearZeroVar(x = bio.012.df,allowParallel = T)

# How many did not meet these thresholds?
length(check.vars)
## 6,624 SNPs did not meet threshold
```

* A total of 6624 SNPs did not meet this threshold, let's take a look at their BCV values

```{r nearzero.stats}
hist(bcv.snps[check.vars])
```

* There is an odd set which have > 3 BCV. Let's subset those out too.

```{r nearzero.bcv}
near.zero.bcv3.snps <- check.vars[which(bcv.snps[check.vars] > 3)]
length(near.zero.bcv3.snps)
```

* A total of 253 SNPs were identified as being removed by the nearzerovar BUT they had BCV > 3. We will keep those for further inspection.

* Do these 253 SNPs show up as significant in the p-adjusted anova scores on the full data?

```{r near.zero.anova}
summary(p.adjust(LGEP.anova.scores)[check.vars])
```

* No, none of these 253 were identified as being associated with the LGEP

* How do the variance of these SNPs look?

```{r near.zero.var}
summary(var.snps[near.zero.bcv3.snps])
```

* The SNP variance's are all pretty small. We will still keep this set and use during prediction.

## Caret pt1: Pressure test multiple models ANOVA < .05

* We have 3 various filters to work with including:

    * Anova scores generated using ONLY LGEP
    
    * BCV values generated using ALL data
    
    * Nearzerovar subset identified with ALL data (BCV > 3)

* In this section, we will run all *faster* models on the training set (LGEP families) and look at which models tend to perform best in terms of `R2` & `rmse`

### Run all *faster* models on LGEP training

* The LGEP and EW are split into train and test using a specific pvalue which will subset SNPs based on their adjusted pval scores

```{r}
pval <- .05
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),which(p.adjust(LGEP.anova.scores) < pval)]
colnames(LGEP.train.dat) <- colnames(fam.012.df[,which(p.adjust(LGEP.anova.scores) < pval)])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),which(p.adjust(LGEP.anova.scores) < pval)]
colnames(EW.test.dat) <- colnames(fam.012.df[,which(p.adjust(LGEP.anova.scores) < pval)])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

* Set up a LOOCV training method for the LGEP and conduct a grid search across parameters of 12 different models.

```
tC <- trainControl(method="LOOCV",allowParallel = T, savePredictions="all",verboseIter = TRUE,search="grid")
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")

t2 <- lapply(algorithmList,function(i){ 
	set.seed(123)
	 t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	 })
save(t2,file="/mnt/anova.05.LGEP.EW.RData")
```

* Extract results from the list and display in table

```{r xtract.sum.mods3,warning=F}
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")
load("/mnt/anova.05.LGEP.EW.RData")
create.data.table()
```

### Performance on EW test set

* Use each model to predict the EW and then display results

```{r EW.pred1}
results.cor <- unlist(lapply(t2,function(i){cor(predict(i,EW.test.dat),EW.test.p)}))
results.rmse <- unlist(lapply(t2,function(i){RMSE(pred = predict(i,EW.test.dat),obs = EW.test.p)}))

plot.mean.model.preds()

plot(results.cor,results.rmse)
which.min(results.rmse)
# model 3 lowest MSE
plot(predict(t2[[3]],EW.test.dat),EW.test.p)

which.max(results.cor)
plot(EW.test.p,predict(t2[[10]],EW.test.dat))
```

* Model 3 

## Caret pt2: Pressure test multiple models ANOVA < .8

* Originally, we just used an adjusted pvalue of 0.05, let's try others..(Note..we can only try so many different SNP sets becuase the total number of SNPs with adjusted p-values less than 0.8 is 339)

* Execute the models

```{r}
pval <- .8
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),which(p.adjust(LGEP.anova.scores) < pval)]
colnames(LGEP.train.dat) <- colnames(fam.012.df[,which(p.adjust(LGEP.anova.scores) < pval)])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),which(p.adjust(LGEP.anova.scores) < pval)]
colnames(EW.test.dat) <- colnames(fam.012.df[,which(p.adjust(LGEP.anova.scores) < pval)])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

```
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")

t2 <- lapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	}
)
save(t2,file = "/mnt/anova.8.LGEP.EW.RData",compress=T)
```

* Extract the results of the models
```{r }
load("/mnt/anova.8.LGEP.EW.RData")
create.data.table()
```

### Performance on EW test set

```{r EW.pred2}
results.cor <- unlist(lapply(t2,function(i){cor(predict(i,EW.test.dat),EW.test.p)}))
results.rmse <- unlist(lapply(t2,function(i){RMSE(pred = predict(i,EW.test.dat),obs = EW.test.p)}))

plot.mean.model.preds()

plot(results.cor,results.rmse)
this.mod <- which.min(results.rmse)
# model 3 lowest MSE
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p)

this.mod <- which.max(results.cor)
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p)
```

## Caret pt3: Near zero var and BCV > 3

* We've tested the anova_score filtering.  Now let's test SNPs which came out in the near zero var 

```{r}
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),near.zero.bcv3.snps]

colnames(LGEP.train.dat) <- colnames(fam.012.df[,near.zero.bcv3.snps])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),near.zero.bcv3.snps]
colnames(EW.test.dat) <- colnames(fam.012.df[,near.zero.bcv3.snps])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

```
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")
t2 <- mclapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	},mc.cores=30)
save(t2,file="/mnt/nearzerovar.bcv3.LGEP.EW.RData",compress=T)
```

* Extract the results of the models

```{r xtract.sum.mods1}
load("/mnt/nearzerovar.bcv3.LGEP.EW.RData")
create.data.table()
```

*Performs badly

## Caret pt4: BCV > 3 var < .2 SNPs

* We've tested the anova_score filtering.  Now let's test those interesting set of SNPs which had a BCV > 3 but variance < .2

* This performs poorly so it's not ran

```
#interesting.bcv.snps <- which(bcv.snps > 3)
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),interesting.bcv.snps]

colnames(LGEP.train.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),interesting.bcv.snps]
colnames(EW.test.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

```
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV")
t2 <- lapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	})
```

* Extract the results of the models

```
create.data.table()
```

*Performs badly

## Caret pt5: BCV > 3 & ANOVA < .05

* Here we take a look at SNPs which had a high BCV and an undajusted anova < .05

```{r}
interesting.bcv.snps <- which(bcv.snps > 3 & LGEP.anova.scores < .05)
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),interesting.bcv.snps]
colnames(LGEP.train.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),interesting.bcv.snps]
colnames(EW.test.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

```
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")
t2 <- mclapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	},mc.cores=30)
save(t2,file="/mnt/bcv3.anova05.LGEP.EW.RData",compress=T)
```

* Extract the results of the models

```{r xtract.sum.mods2}
load("/mnt/bcv3.anova05.LGEP.EW.RData")
create.data.table()
```

* Three models performed relatively well: SVMLinear2, pls, and bagearthGCV

```{r prediction}
results.cor <- unlist(lapply(t2,function(i){cor(predict(i,EW.test.dat),EW.test.p)}))
results.rmse <- unlist(lapply(t2,function(i){RMSE(pred = predict(i,EW.test.dat),obs = EW.test.p)}))

plot.mean.model.preds()

plot(results.cor,results.rmse)
this.mod <- which.min(results.rmse)
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p,xlab = paste0(algorithmList[this.mod]," predictions"))

this.mod <- which.max(results.cor)
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p,xlab = paste0(algorithmList[this.mod]," predictions"))
```

## Caret pt6: BCV > 3 & adjusted ANOVA < .05

* This is pretty interesting because only 5 snps pass that threshold and when we do this prediction we see that only 2 out of the 11 families have predictions which vary instead of being at or close to the mean training BV

```
interesting.bcv.snps <- which(bcv.snps > 3 & p.adjust(LGEP.anova.scores) < .05)
LGEP.train.dat <- fam.012.df[which(fam.phenos$fam_id %in% lgep.fams),interesting.bcv.snps]
colnames(LGEP.train.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
EW.test.dat <- fam.012.df[which(fam.phenos$fam_id %in% ew.fams),interesting.bcv.snps]
colnames(EW.test.dat) <- colnames(fam.012.df[,interesting.bcv.snps])
LGEP.train.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% lgep.fams)]
EW.test.p <- fam.phenos$Volume[which(fam.phenos$fam_id %in% ew.fams)]
```

```
algorithmList <- c("glmnet","pls","bagEarth","bagEarthGCV","cforest","gaussprRadial","knn","penalized","ranger","RRFglobal","svmLinear2","svmRadialCost")
t2 <- mclapply(algorithmList,function(i){ 
	set.seed(123)
	t2 <- train(y=LGEP.train.p, x=LGEP.train.dat, (i), trControl = tC)
	},mc.cores=30)
save(t2,file="/mnt/bcv3.anova05.LGEP.EW.RData",compress=T)
```

* Extract the results of the models

```
load(file="/mnt/bcv3.anova05.LGEP.EW.RData")
create.data.table()
```

* Three models performed relatively well: SVMLinear2, pls, and bagearthGCV

```
results.cor <- unlist(lapply(t2,function(i){cor(predict(i,EW.test.dat),EW.test.p)}))
results.rmse <- unlist(lapply(t2,function(i){RMSE(pred = predict(i,EW.test.dat),obs = EW.test.p)}))

plot.mean.model.preds()

plot(results.cor,df1$x2)
plot(results.cor,results.rmse)
this.mod <- which.min(results.rmse)
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p,xlab = paste0(algorithmList[this.mod]," predictions"))

this.mod <- which.max(results.cor)
plot(predict(t2[[this.mod]],EW.test.dat),EW.test.p,xlab = paste0(algorithmList[this.mod]," predictions"))
```

## Save output

```{r}
save.image("LGEP_EW_pred.caret.snpsV2.RData",compress=T)
```