---
title: "DESeq2 Normalization"
author: "Adam Festa"
date: "May 4, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the txi counts and phenos

* The biological replicate counts and phenotype data were previously prepared for normalization. For more details on the preparation, see Step 2 readme.

```{r load.data}
load("/mnt/media/disk6/ARF/RNASEQ/counts/86kSalmon/Step2_Load_Counts.RData")
names(load.counts)
```

## Create DESeq object

* The object loaded above the txi_object imported using the tximport package and phenotypes.  The [DESeq2 vignette](https://bioconductor.org/packages/3.7/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#transcript-abundance-files-and-tximport-input) describes how to create a dds object using the txi_object.

```{r create.dds, echo=TRUE}
suppressPackageStartupMessages(library(DESeq2))

ddsTxi <- DESeqDataSetFromTximport(load.counts$txi_object,
                                   colData = load.counts$phenos,
                                   design = ~ batch + index_seq + fam_id)
```

## Estimate size factors

* Normalization within DESeq occurs by implementing the function `estimateSizeFactors()`. Once that function finishes we can extract the DESeq2 version of normalized counts.[source link](https://support.bioconductor.org/p/66067/)

```{r normalize}
ddsTxi <- estimateSizeFactors(ddsTxi)
```

## Extract normalized counts, combine with phenos, and save
```{r}
deseq.counts <- counts(ddsTxi,normalized=T)
mypc <- prcomp(t(deseq.counts),scale. = T,center = T)
mypc <- mypc$x
plot(mypc[,1:2])

save(deseq.counts,file = "/mnt/media/disk6/ARF/RNASEQ/counts/86kSalmon/Step3_DESEQ_Counts.RData",compress=T)
```