---
title: "DESeq2 + LMM No ide term"
author: "Adam Festa"
date: "May 4, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the txi counts and phenos ####

* The biological replicate counts and phenotype data were previously prepared for normalization. For more details on the preparation, see Step 2 readme.

```{r load.data}
load("/media/titan/disk6/ARF/RNASEQ/counts/86kSalmon/Step2_Load_Counts.RData")
names(load.counts)
```

## Create DESeq object ####

* The object loaded above the txi_object imported using the tximport package and phenotypes.  The [DESeq2 vignette](https://bioconductor.org/packages/3.7/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#transcript-abundance-files-and-tximport-input) describes how to create a dds object using the txi_object.

```{r create.dds, echo=TRUE}
suppressPackageStartupMessages(library(DESeq2))

ddsTxi <- DESeqDataSetFromTximport(load.counts$txi_object,
                                   colData = load.counts$phenos,
                                   design = ~ batch + index_seq + fam_id)
```

## Estimate size factors ####

* Normalization within DESeq occurs by implementing the function `estimateSizeFactors()`. Once that function finishes we can extract the DESeq2 version of normalized counts.[source link](https://support.bioconductor.org/p/66067/)

```{r normalize}
ddsTxi <- estimateSizeFactors(ddsTxi)
```

## Extract normalized counts ####
```{r}
deseq.counts <- counts(ddsTxi,normalized=T)
```

## Obtain pedigree information ####

* Generate a csv file of unique family names to obtain pedigree information by subsetting the unique IDs from the phenotype data frame 

```{r}
unique.fams <- unique(unlist(strsplit(x = unique(as.character(load.counts$phenos$fam_id)) , split = "x")))
unique.fams <- data.frame("ID"=unique.fams)
#Write out a CSV file which can be uploaded to TIPRoot to access pedigree
    #write_csv(unique.fams,col_names = T,path ="/media/disk6/ARF/RNASEQ/Breeding-Value-Prediction/disk6directory/resources/phenos/ufams.v2.csv")
```

* Steps to Generate Ped:
    * Accessed tip root
    * pedigree infomration --> pedigree and relatives
    *  upload csv written up and generate ancestors
    *  download file and put in /resources/pedigree/
    *  Open file and delete the comment column

## Load pedigree ####

```{r}
library(readxl)
  pedigree <-read_xlsx("/media/titan/disk6/ARF/RNASEQ/Breeding-Value-Prediction/disk6directory/resources/pedigree/pedigree_2018_04_19_20_36_15.xlsx")
```

* Note that the pedigree returned from tiproot does not contain any WG families

```{r}
  as.character(unique.fams[,1])[which(!as.character(unique.fams[,1]) %in% as.character(pedigree$Id))]
```
* We can add those and full sib crosses next

## Add missing OP families to the pedigree (WG fams) ####
```{r}
    add.fams <- as.character(unique.fams[,1])[which(!as.character(unique.fams[,1]) %in% as.character(pedigree$Id))]
  #These are the families which are missing
    head(add.fams)
  #Create data frame for families which are missing from the pedigree
    missing.ped <- data.frame("Id"=add.fams,"Parent 1"=NA,"Parent 2"=NA,"County"=NA,"State"=rep("WG",length(add.fams)))
    colnames(missing.ped) <- c("Id","Parent 1","Parent 2","County","State")
  #Rbind the missing pedigree to the one obtained from tiproot
    pedigree <- rbind(pedigree,missing.ped)
    rm(unique.fams,add.fams,missing.ped)
  #Note**Some families are missing a State while most are WG one is N11011..what's up with that?
    pedigree$Id[which(is.na(pedigree$State) == T)]
```

## Create progeny pedigree

```{r}
  prog.pedigree <- unique(load.counts$phenos[,c("animal_id","p1","p2")])
  colnames(prog.pedigree) <- c("animal","sire","dam")
```

## Add missing FS families to the pedigree

```{r}
  # Get list of unique family Id's from experiment
    unique.fams <- unique(as.character(load.counts$phenos$fam_id))
  # Identify FS crosses missing from pedigree
    add.fams <- as.character(unique.fams)[which(!as.character(unique.fams) %in% as.character(pedigree$Id))]
    head(add.fams)
  #Parent1 will be the first parent before the "x" & Parent2 will be the second
    fs.par1 <- unlist(lapply(strsplit(add.fams,split = "x"),function(x) x[1]))
    fs.par2 <- unlist(lapply(strsplit(add.fams,split = "x"),function(x) x[2])) 
  #Fill in families to pedigree
    missing.ped <- data.frame("Id"=add.fams,"Parent 1"=fs.par1,"Parent 2"=fs.par2,"County"=NA,"State"=rep("FS",length(add.fams)))
    colnames(missing.ped) <- c("Id","Parent 1","Parent 2","County","State")
  #Rbind the missing pedigree to the one obtained from tiproot
    pedigree <- rbind(pedigree,missing.ped)
    rm(unique.fams,add.fams,fs.par1,fs.par2,missing.ped)
  #Note**Some FS families will be a mix of the state they came from.
```

## Complete pedigree 
```{r}
  # Create A inverse from pedigree
    head(pedigree)
    asreml.ped <-data.frame(pedigree[,1:3])
    colnames(asreml.ped) <- c("animal","dam","sire")
    asreml.ped <- within(asreml.ped,{animal <- factor(animal); dam <- factor(dam); sire <- factor(sire)})
library(asreml)
    ainv <- asreml.Ainverse(pedigree = asreml.ped)$ginv
```


## Normalize all of the data together
```{r}
    pre.norm.counts <- data.frame(apply(deseq.counts,2,as.integer))
      colnames(pre.norm.counts) <- load.counts$phenos$animal_id
    pre.norm.length <- load.counts$txi_object$length
      colnames(pre.norm.length) <- load.counts$phenos$animal_id
    pre.norm.var <- load.counts$txi_object$variance
      colnames(pre.norm.var) <- load.counts$phenos$animal_id
    expt.dat.tmp <- data.frame(load.counts$phenos[,c("animal_id","fam_id","index_seq","batch")],stringsAsFactors = T)
```

```{r}
pt <- proc.time()
    asreml.de.counts.bio.noide <- mclapply(1:nrow(pre.norm.counts),function(each.txpt){
      #each.txpt=323
    tip.subset <- data.frame(cbind("txpt"=unlist(pre.norm.counts[each.txpt,]),expt.dat.tmp,"length" = pre.norm.length[each.txpt,]))

    tip.subset <- within(tip.subset,{txpt <- log2(as.integer(txpt)+1); 
      animal_id <- factor(animal_id); 
      batch <- factor(batch);
      fam_id <- factor(fam_id); 
      true.index <- factor(index_seq); 
      length <- as.numeric(length)})

    tmp1 <- asreml(fixed= txpt ~1 + batch  + true.index, random = ~ ped(fam_id,var=T,init=1),
               ginverse=list(fam_id=ainv),data=tip.subset,maxiter=50,trace=F,
               family=asreml.gaussian(link = "identity"))

    edit.names <- gsub(pattern = "fam_id, var = T)_",replacement = "",x = names(tmp1$coefficients$random))
      edit.names <- gsub("ped\\(", "", edit.names)
    my.fams <- which(edit.names %in% unique(as.character(load.counts$phenos$fam_id)))
    my.counts <- tmp1$coefficients$random[my.fams] + tmp1$coefficients$fixed['(Intercept)']
      my.counts},mc.cores=23)
proc.time() - pt
```

153 seconds for 20K txpts (approx 2.5 minutes * 4 sets of 20k = ~10 minutes for normalization)


## Combine count output and save 
```{r}
asreml.de.counts.bio.noide <- do.call(cbind,asreml.de.counts.bio.noide)
rm.zero.var <- (which(apply(asreml.de.counts.bio.noide,2,function(x) {var(x)}) == 0))
asreml.de.counts.bio.noide <- asreml.de.counts.bio.noide[,-rm.zero.var]
mypc <- prcomp(2^asreml.de.counts.bio.noide,scale. = T,center = T)
mypc <- mypc$x
plot(mypc[,1:2])

edit.names <- gsub(pattern = "fam_id, var = T)_",replacement = "",x = rownames(asreml.de.counts.bio.noide))
      edit.names <- gsub("ped\\(", "", edit.names)
    
rownames(asreml.de.counts.bio.noide) <- edit.names
colnames(asreml.de.counts.bio.noide) <- rownames(pre.norm.counts)[-rm.zero.var]
asreml.de.counts.bio.noide <- 2^asreml.de.counts.bio.noide
save(asreml.de.counts.bio.noide,file = "../Step3_DE_LMM_bio.noide_Counts.RData",compress=T)    
```  
    